Using Sorghum as an outgroup to find syntenic regions in Maize: regenerating

### 2017-09-22 SubGenome Analysis.Rmd
Analysis of Subgenomes by expression data

@@

@@

>>>>

Expression data was retrieved from [NCBI: GSE50191](https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE50191) and was reported in [Walley 2016](http://science.sciencemag.org/content/353/6301/814.full).  This set includes 69 experiments (3 biological replicates, 23 tissues) for 62,547 mRNAs aligned to Zea mays B73 RefGen_v2 5a WGS annotation set.

@@

Paired t-test suggests that there is no difference in means between FPKM of homeologs in sub1 vs. sub2 (p-val > 0.05, so we accept null hypothesis of equal means):
```{r subgenome_fpkm_ttest, echo=FALSE}
t.test(x=expressedPairs$FPKM_mean1, y=expressedPairs$FPKM_mean2, paired = TRUE)
```

@@

I am curious if our definition of sub 1 and sub 2 (i.e. the larger syntenic block is defined as sub 1) is really the best way to decide which genes are in sub 1 and sub 2. It seems to me that if one could resonably expect that, given 2 identically duplicate genes, if there was no benefit to having duplicated function then it would be a matter of random chance which of the two duplicate genes would loose function over time.  It also seems to me that the quickest way to "degrade" an unnecessary gene (evolutionarily speaking) is to reduce expression.  So what happens if we pick the higher expressing gene to be in sub 1 and the lower expressing gene to be in sub 2?  Not surprising, we can find differences in populations between high and low expressing pairs.  I wonder if this holds across different maize lines?

```{r subgenome_expression_high_v_low, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=4.5, fig.height=4, fig.align='center'}
data <- expressedPairs
data$LowExpressingGene <- apply(data[,c(3,4)], 1, FUN=min)
data$HighExpressingGene <- apply(data[,c(3,4)], 1, FUN=max)
data <- data %>% gather(key = type, value = value, 5:6)
ggplot(data, aes(type, log10(value))) +
  geom_boxplot() +
  labs(y = "log10(FPKM)",
       x = "Types",
       title = "Comparison of FPKM Expression between\nGenes in high express vs. low express groups"
  )
```

```{r subgenome_expression_dominance, echo=FALSE, warning=FALSE, message=FALSE, results='hide', fig.width=10, fig.height=4, fig.align='center'}
plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 1, TRUE)
# plot <- graphGenePairExpressionsByExperiement(maize.expression.all, homeologs.pairs, 2:12, 4, FALSE)
plot(plot)
```
